---
title: "Lab08 Classwork"
author: "Priya Jindal"
date: "10/24/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```
```{r}
k <- kmeans(x, centers = 2, nstart = 20)
k

```

How many points in each cluster? -> 30
What component is your result object details
    -cluster size -> "K-means clustering with 2 clusters of sizes 30, 30" k$size
    -cluster membership -> clustering vector, which point is in what cluster 
    -cluster center -> cluster means vectors 
    
```{r}
k$size
k$cluster #clustering vector
k$centers
```

```{r}
plot(x, col = k$cluster)
points(k$centers, col = "green", pch=15)
```
```{r}
dist_matrix <- dist(x)
hc <- hclust(d=dist_matrix)
plot(hc)
cutree(hc, k=4)
```

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```

```{r}
hc <- hclust(dist(x))
plot(hc)
abline(h=2, col="red")
grps <- cutree(hc, h = 2)
```
```{r}
plot(x, col=grps)
table(grps, col)

```
```{r}

mydata <- read.csv("https://tinyurl.com/expression-CSV", row.names =1)

head(mydata)
#x <- read.csv("data/UK_foods.csv")
```
#How many genes are in the dataset?
```{r}
dim(mydata)
nrow(mydata)
```

#Lets do PCA 
```{r}
## lets do PCA
pca <- prcomp(t(mydata), scale=TRUE)
## See what is returned by the prcomp() function

## A basic PC1 vs PC2 2-D plot 
plot(pca$x[,1], pca$x[,2]) 

## Precent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1) 
pca.var.per

attributes(pca)
# $names
#[1] "sdev" "rotation" "center" "scale" "x"
#
# $class
#[1] "prcomp"
```
```{r}
barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")
```
```{r}
## A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)")) 
```


# Classwork Hands-On Section s
```{r}

x <- read.csv("data/UK_foods.csv")
x

```

Q1: How many rows/columns are in your new data frame names x
```{r}
dim(x)
```

Fix the row names set up 
```{r}
rownames(x) <- x[,1]
x <- x[,-1]
head(x)
```
```{r}
x <- read.csv("data/UK_foods.csv", row.names=1)
head(x)

```
Q2: Which approach to solwing row name problem is better
x <- read.csv("data/UK_foods.csv", row.names=1)

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```
Q3: if we leave out the optional param beside then we can get the other type of bar graph 
```{r}
barplot(as.matrix(x), col=rainbow(nrow(x)))
```

```{r}
pairs(x, col=rainbow(10), pch=16)
```
Q4: If the points lie in the diagonal that means those 2 countries's trends are positvelly correlating. 

#PCA practice 
```{r}
pca <- prcomp(t(x))
summary(pca)
```

```{r}
#Plot Pc1 vs pc2 
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x))
```

